\color{navyblue} Difference-in-Differences Estimation with Spatial Spillovers

### 2021-04-23 latex/paper/Spillover.tex
Draft 4/16

\title{\color{navyblue} Difference-in-Differences Estimation with Spatial Spillover}
% \subtitle{Subtitle}
\author{\normalsize Kyle Butts\\{\footnotesize Univ. of Colorado, Boulder}}
\date{\footnotesize\today}

@@

Empirical work in economics often considers settings where treatment is assigned to groups of units, but the effect of these treatments spillover onto ‘close’ units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic
distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} For example, treatment is commonly assigned to geographic boundaries and spillovers occur from people crossing borders to access treatment (i.e. imperfect compliance) or from general equilibrium effects influencing neighboring areas. Despite the problem of spillovers being common across many settings, \citet{Berg_Streitz_2019} document that little empirical analysis directly controls for spillovers when estimating the `direct' effect of treatment on the treated units. The term `direct' effect can refer to either the average treatment effect on the treated or the intent to treat effect.\footnote{In empirical applications, the `direct` effect is sometimes called the `partial equilibrium' effect or the `local' effect.} The central theoretical result of my paper is that in the presence of spillovers, the standard difference-in-differences estimate identifies the direct effect of treatment (the estimand of interest) plus two additional bias terms resulting from the spillovers.

The intuition for the two bias terms resulting from spillovers is as follows. First, untreated units that are `close' to treated units experience effects of treatment and therefore these `control' units fail to identify the counterfactual trend. When estimating by difference-in-differences, the spillover onto the `close' control units is averaged into the untreated units' change in outcomes. In this case, the spillover is subtracted from the estimated treatment effect and biases the estimate in the opposite sign of the spillover effect. For example, consider trying to estimate employment effects of a factory opening in a county. Agglomeration economies suggest that neighboring counties would also benefit from knowledge spillovers and improved access to supply chain which would potentially increase economic activity \citep{Duranton_Puga_2003}. The treatment effect estimate is negatively biased because the change in outcome in neighboring counties is higher than it would be absent treatment. Researchers would therefore underestimate the positive benefits of treatment.\footnote{If the policy maker had a welfare function that included both treated and neighboring counties, they would fail to count the positive benefits of the control units and they would underestimate the positive benefits on the treated units. Difference-in-differences estimation would therefore result in double undercounting.}

% @TODO: Improve examples to better build intuition

Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. The spillover onto other treated units is averaged into the treated units' change in outcomes. Therefore the bias of treatment effect is the same sign as the sign of the spillover. Continuing with our example, two factory openings in neighboring counties might cause the benefit of each individual factory to increase due to agglomeration forces. The estimated treatment effect will count the `direct' effect of the factory opening as well as the benefit of the agglomeration economies from nearby factory openings. Therefore the treatment effect will be positively biased due to the positive spillover onto also treated units.

I then use my framework to evaluate commonly used empirical strategies found in applied work and provide practical recommendations for researchers. The most common approach is to remove potentially contaminated control units from the sample in order to remove spillover effects on control units \citep{Berg_Streitz_2019}. This approach is not recommended for two reasons. First, the removal of too few control units will leave bias and removal of too many control units will lower precision of the treatment effect estimate. This results in a bias-variance trade-off that other approaches do not share. Second, this method does not remove spillover effects on treated units and will leave this bias term in the difference-in-differences estimate. Another common approach is to include either an indicator or a set of rings around treated units. This approach will remove all bias from the spillover effects on control units provided that all control units with spillover effects are included in the indicator. Similar to the previous approach, spillover effects on also treated units still add bias to estimates of the direct effect.

Therefore, I propose an estimation strategy that will remove both sources of bias with very minimal assumptions on the structure of spillovers. To remove \textit{all bias} from the direct effect estimate, a research needs to just include an indicator for being close to a treated unit interacted with treatment status, so long as the indicator captures all units affected by spillovers. In my decomposition of the treatment effect estimate, the bias terms are the \textit{average} spillover effects onto treated and control units. An idicator variable interacted with treatment status therefore can estimate these two averages and remove them from the difference-in-differences estimate. Importantly, my method does not require researchers to make any assumptions about how spillover effects propagate across space. The only assumption required is the maximum distance from treated units spillovers can occur.\footnote{Even if the maximum distance is not large enough, since treatment effects typically decay over distance, most of the bias will still be removed. However, including too many units will increase the variance in the estimates for average spillover effects.}

Since spillover effects are often important causal effects themselves, I next turn to how to estimate spillovers directly. I show with evidence from Monte Carlo simulations, that commonly used semi-parametric estimation strategies capture spillovers well (in a mean square prediction error sense). This involves creating a set of distance bins from treated units (e.g. being 0-20 miles, 20-40 miles, 40-60 miles from treated unit) and interacting them with a treatment indicator.\footnote{The choice of the distance bins depends on the economic context and in particular the source of the spillovers.
% TODO: Reference Butts and Clarke
} The key consideration required by a researcher is whether the size of spillover effects are additive in the number of nearby treated units or not. For non-additive spillovers, I recommend a set of mutually-exclusive indicators which measures which distance bin that the closest treated unit falls within for a given observation. For additive spillovers, I recommend using the number of treated units within each distance bin for a given observation.\footnote{While the non-additive version of ring indicators will remove all bias from the direct effect of treatment even under misspecification, this is not true of additive rings. Therefore, researchers may want to estimate the direct effect of treatment using non-additive rings and then use additive rings to estimate spillover effects.}

I turn to an empirical application in the literature in urban economics on local place-based policies in Section \ref{sec:tva}. Using structural urban models, \citet{Kline_Moretti_2014b} show that the `local' or partial equilibrium effects of a place-based policy are offset in part by indirect or general equilibrium effects in non-treated locations. In particular, their model implies these effects are declining in distance. I contribute by showing that because of these general equilibrium effects, estimation of local effects are often misestimated by not controlling for the spillover effects in their empirical specification.

As an example, I revisit the analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. The Tennessee Valley Authority was a large scale New Deal program that created many new dams for flood-protection and navigation. The construction of large-scale dams lowered the cost of power for industrial firms giving the region an industrial boost \citep{Kitchens_2014}. The scale of federal investment in the region was large and the pro-manufacturing benefits likely spread further than the Authority's boundary due to the electrification infrastucture and agglomeration economies \citep{Severnini_2014}. I show that estimation by difference-in-differences fails to account for these spillovers and for different outcome variable it underestimates and overestimates the local effect, depending on the sign of the spillovers.

Following this empirical application, I discuss how my framework fits into a larger discussion on identification strategies with place-based policies. I revisit conflicting results about United States' federal Empowerment Zones, a program that creates incentives for businesses to locate in high-poverty neighborhoods. \citet{Busso_Gregory_Kline_2013} use Census Tracts from qualified but ultimately rejected applications that are typically far away from accepted Zones as a comparison group and find that significant reductions in poverty. This empirical strategy is based on the idea that since they qualified for the program they would be on similar trends without being contaminated by spillover effects. \citet{Neumark_Kolko_2010}, on the other hand, use census tracts within 1,000 feet of the Zone and find no statistically significant effects. This empirical strategy assumes that the borders are drawn somewhat randomly and therefore being just outside is as good as random. Although, using close units is potentially problematic as they can experience spillover effects from the Zones. My framework can explain the differences in findings if there are positive spillover effects onto census tracts just outside Empowerment Zones.

@@

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} presents Monte Carlo simulations to illlustrate the bias result and to evaluate currently used solutions in the literature. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}.

@@

Following the canonical difference-in-difference framework, there is a time $t_0$ where treatment turns on and remains on afterwards. This framework is extended to staggered treatment adoption in section \ref{sec:event_study}. Potential outcomes for unit $i$ at time $t$ are a function of own treatment-status $D_i$ and, departing from the standard potential outcomes framework, of a function of the entire vector of treatment assignments $h(\vec{D}, i)$ where $\vec{D} \in \{0,1 \}^n$ denotes the $n$-dimensional vector of all unit treatments. The potential outcomes are denoted $Y_{i,t}(D_i, h(\vec{D}, i)$). The function $h(\vec{D}, i)$ is referred to as an `exposure mapping' and is a non-negative scalar- or vector- valued function. The exposure mapping measures the intensity at which unit $i$ is affected by spatial spillovers. When unit $i$ is sufficiently `far' away, it has no exposure to spatial spillovers and $h(\vec{D}, i) = \vec{0}$. The exposure mapping formalizes the SUTVA violation in that potential outcomes are affected by other unit's treatment assignment as summarized by $h(\vec{D}, i)$. To help better understand the exposure mapping function, I give three examples of $h(\vec{D}, i)$ that are commonly used in the literature.

\begin{example}
    First, $h(\vec{D}, i)$ can be a 0/1 variable that equals one only if there is a treated unit within $\bar{d}$-miles of unit $i$.\footnote{Similarly a dummy for counties that share borders is commonly used.} Let $d(i,j)$ be a geographic distance measure which tells the distance unit $i$ is from unit $j$.  In this case \begin{equation}\label{eq:h_within}
        h(\vec{D}, i) = \max_{i \neq j} D_j * 1[ d(i,j) < \bar{d} ] 
    \end{equation}
    This exposure mapping is useful when it is assumed that spillovers do not decay over distance until $\bar{d}$ and the number of neighboring units treated do not impact the intensity of spillovers. For example, this exposure mapping likely applies in the context of new library creation \citep{Berkes_Nencka_2020}. In this case the distance $\bar{d}$ would be the maximum distance people would travel to a nearby library. Access to a neighboring town library is binary in nature, so it does not matter whether you can access 1 or more nearby libraries, so the binary exposure mapping is a good approximation to spillovers in this context.  
\end{example}

\begin{example}
    Second, $h(\vec{D}, i)$ can be a function that equals the number of units treated within distance $\bar{d}$, i.e. \begin{equation}\label{eq:h_within_additive}
        h(\vec{D}, i) = \sum_{j} D_j * 1[ d(i,j) < \bar{d} ].
    \end{equation}
    This exposure mapping is no longer binary, so the intensity of spillovers depend on the number of nearby units treated. In the context of large store openings, this exposure mapping captures the additive nature of agglomeration economies. As more nearby counties receive new stores, the agglomeration forces increase \citep{Basker_2005}.
\end{example}

\begin{example}
    Last, is the spatial decay function where exposure decreases with distance. In this case, spillover intensity is the sum across all treated observations' decay term, i.e. \begin{equation}\label{eq:h_decay}
        h(\vec{D}, i) = \sum_{j \neq i} D_j e^{-\alpha d(i,j)}.
    \end{equation} 
    This exposure mapping allows for the intensity of spillovers to depend on distance to treatment and also is additive in the number of nearby units treated.\footnote{However, this specification assumes that all units are affected by all other units. This creates problems with inference because it implies potential correlation between all units' error terms. For this reason, this function often is summed over only the $k$-nearest neighbors or over units within $\bar{d}$ miles.} The speed of spillover decay over distance depends on the parameter $\alpha$ that can be calibrated by the researcher or estimated using non-linear least squares. In the literature on R\&D investment, for example, \citet{Keller_2002} uses a modified version of this exposure mapping where $D_j$ is country $j$'s R\&D expenditure. 
\end{example}

After chosing an exposure mapping that makes sense in the economic context, a researcher must also specify the functional form of the potential outcomes. Typically, the spillover function enters into the regression linearly and potentially the coefficient is allowed to differ by treatment status.

@@

With the potential outcomes defined, I now formalize what is meant by `spatial spillovers'. I define `spillover onto control units' as: \[
    Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}).
\] 
The spillover measures the difference in non-treated potential outcomes between being exposed at intensity $h(\vec{D}, i)$ and not being exposed. Then, the average spillover effect onto control units averages over potential heterogeneity in the effect size of spillovers and over heterogeneity in exposure intensity $h(\vec{D}, i)$: \[
    \tau_{\text{spill,control}} \equiv \mathbb{E} \left[ Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 0 \right].
\]

To emphasize, the average spillover effect onto conrol units averages over each control unit's exposure mapping. For example, assume the potential outcomes is additively linear in $D_i$ and $h(\vec{D}, i)$ and that the coefficient $\beta_{\text{spill,control}}$ measures the effect of $h(\vec{D}, i)$ on outcome $Y$ among control units. Then an individual control unit's spillover effect is $\beta_{\text{spill,control}} \ h(\vec{D}, i)$. The average spillover effect onto control unit would therefore be $\tau_{\text{spill,control}} = \beta_{\text{spill,control}} * \mathbb{E} \left[ h(\vec{D}, i)\right]$, i.e. the average over all control units exposure mapping.

Similarly, we define the average spillover effect onto also treated units as: \[ 
    \tau_{\text{spill,treated}} \equiv \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0) \ \vert \ D_i = 1 \right].
\]

It is important to clarify what I am assuming is the estimand of interest researchers would like to estimate when using difference-in-differences. I assume that what the `average treatment effect' is trying to measure in this context is what I will call the `direct effect of treatment': \[
    \tau_{\text{direct}} = \mathbb{E} \left[ Y_{it}(1, \vec{0}) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 1 \right],
\] 
which measures the effect of being treated in the absence of exposure to spillovers. This effect does not include the general equilibrium effects of multiple treated areas near one-another.

My definition of the direct effect of treatment differs from \citet{Sävje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(0, h(\vec{D}, i)) \ \vert \ D_i = 1 \right],
\] 
where the expectation is over individuals and their exposures. I prefer the former because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. For example, a county deciding whether to implement a policy will want to consider the treatment effect plus the spillover at their current level of exposure. The estimand proposed by \citet{Sävje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at different levels.\footnote{Estimation of the direct effect and spillover effect can be combined to estimate the ATE as defined by \citet{Sävje_Aronow_Hudgens_2019}.}

@@

In this section, I identify the two sources of bias in difference-in-differences estimation of the direct effect of treatment. To estimate the `direct effect of treatment', researchers estimate the canonical two-way fixed effects model, 
\begin{equation}\label{eq:twfe}    
    y_{it} = \tau D_{it} + \mu_i + \mu_t + \epsilon_{it}.
\end{equation}

The estimator $\hat{\tau}$ is a biased estimate for $\tau_{\text{direct}}$ in the presence of spillovers. To show this, I first present the equivalent to the parallel counterfactual trends assumption in the context of the new potential outcome framework.

\begin{assumption}[Parallel Counterfactual Trends]\label{eq:parallel}
    \[ 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 1 \right] = 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 0 \right]
    \]
\end{assumption}
This assumption states that in the absence of treatment and with zero exposure (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status. This generalizes to the classic parallel counterfactual trends when SUTVA is satisfied because then every unit has zero exposure.

Given that Assumption \ref{eq:parallel} holds, the estimate $\hat{\tau}$ from (\ref{eq:twfe}) can be decomposed as the direct effect and the two sources of spillover bias. The proof is given in Appendix \ref{sec:proofs}.

>>>>

\title{\color{navyblue} Difference-in-Differences Estimation with Spatial Spillovers}
\author{\normalsize Kyle Butts\\{\footnotesize Univ. of Colorado, Boulder}}
\date{\footnotesize\today}

@@

Empirical work often considers settings where treatment is assigned to groups of units by geographic boundaries, but the effect of these treatments spillover onto `nearby' units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} In this setting, there are two types of spillover effects. First, there can be spillover effects onto control units. For example, individuals in control areas can travel to the treated areas and receive treatment (e.g. a hospital opening serves nearby residents) or changes in the economy in a treated area can have feedback loops that affect nearby economies (e.g. a new factory opening increases service sector spending). Second, there can be spillover effects onto other treated units. For example, forces of agglomeration can increase effects as treatment concentration increases (e.g. knowledge-sharing between workers), forces of congestion or market competition can decrease the effects of treatment (e.g. bidding up of wages cause employment effects of factory openings to be mitigated), and treatment effectiveness may be improved by multiple jurisidictions learning from one another.

In this paper, I introduce a potential outcome framework that directly models spillover effects. Using this framework, the central theoretical result of my paper is that in the presence of spillovers, the standard difference-in-differences estimate identifies the direct effect of treatment (the estimand of interest) plus two additional bias terms resulting from the spillovers.\footnote{The term `direct' effect can refer to either the average treatment effect on the treated or the intent to treat effect. In empirical applications, the `direct` effect is sometimes called the `partial equilibrium' effect or the `local' effect.} The intuition for why the two forms of spillovers cause bias is as follows.

First, untreated units that are `close' to treated units experience effects of treatment and therefore these `control' units fail to identify the counterfactual trend. When estimating by difference-in-differences, the spillover onto the `close' control units is averaged into the untreated units' change in outcomes. In this case, the spillover is subtracted from the estimated treatment effect and biases the estimate in the opposite sign of the spillover effect.

For example, consider trying to estimate employment effects of a factory opening in a county. Agglomeration economies suggest that neighboring counties would also benefit from knowledge spillovers and improved access to supply chain which would potentially increase economic activity \citep{Duranton_Puga_2003}. The treatment effect estimate is negatively biased because the change in outcome in neighboring counties is higher than it would be absent treatment. Researchers would therefore underestimate the positive benefits of treatment.\footnote{If the policy maker had a welfare function that included both treated and neighboring counties, they would fail to count the positive benefits of the control units and they would underestimate the positive benefits on the treated units. Difference-in-differences estimation would therefore result in double undercounting.}

% Despite the problem of spillovers being common across many settings, \citet{Berg_Streitz_2019} document that little empirical analysis directly controls for spillovers when estimating the `direct' effect of treatment on the treated units.

Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units.The spillover onto other treated units is averaged into the treated units' change in outcomes. Therefore the bias of treatment effect is the same sign as the sign of the spillover. Continuing with our example, two factory openings in neighboring counties might cause the benefit of each individual factory to increase due to agglomeration forces. The estimated treatment effect will count the `direct' effect of the factory opening as well as the benefit of the agglomeration economies from nearby factory openings. Therefore the treatment effect will be positively biased due to the positive spillover onto also treated units.

I then use my potential outcome framework to evaluate commonly used empirical strategies found in applied work and provide practical recommendations for researchers. The most common approach is to remove potentially contaminated control units from the sample in order to remove spillover effects on control units \citep{Berg_Streitz_2019}. This approach is not recommended for two reasons. First, the removal of too few control units will leave bias and removal of too many will lower precision of the treatment effect estimate. This results in a bias-variance trade-off that other approaches do not share. Second, this method does not remove spillover effects on treated units and will leave this bias term in the difference-in-differences estimate. Another common approach is to include either an indicator or a set of rings around treated units. This approach will remove all bias from the spillover effects on control units provided that all control units with spillover effects are included in the indicator. Similar to the previous approach, spillover effects on also treated units still add bias to estimates of the direct effect.

Therefore, I propose an estimation strategy that will remove both sources of bias with very minimal assumptions on the structure of spillovers. To remove \textit{all bias} from the direct effect estimate, a researcher needs to just include an indicator for being close to a treated unit interacted with treatment status, so long as the indicator captures all units affected by spillovers. In my decomposition of the treatment effect estimate, the bias terms are the \textit{average} spillover effects onto treated and control units. An indicator variable interacted with treatment status therefore can estimate these two averages and remove them from the difference-in-differences estimate. Importantly, my method does not require researchers to make any assumptions about how spillover effects propagate across space. The only assumption required is the maximum distance from treated units spillovers can occur.\footnote{Even if the maximum distance is not large enough, since treatment effects typically decay over distance, most of the bias will still be removed. Including too many units in the indicator will increase the variance in the estimates for average spillover effects as the other control units identify the parallel counterfactual unit.}

Since spillover effects are often important causal effects themselves, I next turn to how to estimate spillovers directly. I show with evidence from Monte Carlo simulations, that commonly used semi-parametric estimation strategies capture spillovers well (in a mean square prediction error sense). This involves creating a set of distance bins from treated units (e.g. being 0-20 miles, 20-40 miles, 40-60 miles from treated unit) and interacting them with a treatment indicator.\footnote{The choice of the distance bins depends on the economic context and in particular the source of the spillovers.} The key consideration required by a researcher is whether the size of spillover effects are additive in the number of nearby treated units or not. For non-additive spillovers, I recommend a set of mutually-exclusive indicators which measures which distance bin that the \textit{closest} treated unit falls within for a given observation. For additive spillovers, I recommend using the \textit{number of} treated units within each distance bin for a given observation.\footnote{While the non-additive version of ring indicators will remove all bias from the direct effect of treatment even under misspecification, this is not true of additive rings. Therefore, researchers may want to estimate the direct effect of treatment using non-additive rings and then use additive rings to estimate spillover effects.}

To show the importance of considering spillover effects, I revisit analyses of place-based policies in urban economics in Section \ref{sec:tva}. I revisit the analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. The Tennessee Valley Authority was a large scale New Deal program that created many new dams for flood-protection and navigation. The construction of large-scale dams lowered the cost of power for industrial firms giving the region an industrial boost \citep{Kitchens_2014}. The scale of federal investment in the region was large and the pro-manufacturing benefits likely spread further than the Authority's boundary due to the electrification infrastucture and agglomeration economies \citep{Severnini_2014}. I show that estimation by difference-in-differences fails to account for these spillovers and therefore forms biased estimates of the local effect of the Tennessee Valley Authority.

Following this empirical application, I discuss how my framework fits into a larger discussion on identification strategies with place-based policies. I revisit conflicting results about United States' federal Empowerment Zones, a program that creates incentives for businesses to locate in high-poverty neighborhoods. \citet{Busso_Gregory_Kline_2013} use Census Tracts from qualified but ultimately rejected applications that are typically far away from accepted Zones as a comparison group and find that significant reductions in poverty. This empirical strategy is based on the idea that since they also qualified for the program, these comparison units would be on parallel trends without being contaminated by spillover effects from treatment. \citet{Neumark_Kolko_2010}, on the other hand, use census tracts within 1,000 feet of the Zone and find no statistically significant effects. This empirical strategy assumes that the borders are drawn somewhat randomly and therefore being just outside is as good as random. Although, using close units is potentially problematic as they can experience spillover effects from the Zones. My framework can explain the differences in findings if there are positive spillover effects onto census tracts just outside Empowerment Zones.

@@

I demonstrate the method by revisiting the analysis by \citet{Bailey_Goodman_Bacon_2015} of Community Health Centers which provided low-cost primary care to impoverished areas. They find that the health centers significantly lowered the mortality rate in treated counties. I find that in this context, since the mortality declines caused by the health centers was due to primary care, spillover effects are near-zero. This result shows the importance of accessability concerns when deciding the location of the health centers.

% ------------------------------------------------------------------------------
\subsection{Relation to the Literature} 
% ------------------------------------------------------------------------------

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} focuses on estimation of the direct and spillover effects of treatment. I evaluate currently used solutions in the literature and present a set of recommendations to more robust estimation strategies. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}. I apply these methods in evaluating spillover effects of U.S. Community Health Centers in Section \ref{sec:chc}.

@@

Following the canonical difference-in-differences framework, there is a time $t_0$ where treatment turns on and remains on afterwards.\footnote{This framework is extended to staggered treatment adoption in Section \ref{sec:event_study}.} Potential outcomes for unit $i$ at time $t$ are a function of own treatment-status $D_i$ and, departing from the standard potential outcomes framework, of a function of the entire vector of treatment assignments $h(\vec{D}, i)$ where $\vec{D} \in \{0,1 \}^n$ denotes the $n$-dimensional vector of all unit treatments. The potential outcomes are denoted $Y_{i,t}(D_i, h(\vec{D}, i)$). The function $h(\vec{D}, i)$ is referred to as an `exposure mapping' and is a non-negative scalar- or vector- valued function. The exposure mapping measures the intensity at which unit $i$ is affected by spatial spillovers. When unit $i$ is sufficiently `far' away, it has no exposure to spatial spillovers and $h(\vec{D}, i) = \vec{0}$. The exposure mapping formalizes the SUTVA violation by summarizing how outcomes are affected by other unit's treatment assignment. To help better understand the exposure mapping function, I give three examples that are commonly used in the literature.

\begin{example}
    First, $h(\vec{D}, i)$ can be an indicator variable that equals one only if there is a treated unit within $\bar{d}$ miles of unit $i$.\footnote{Similarly a dummy for counties that share contiguous borders is commonly used.} Let $d(i,j)$ be a distance measure which tells the distance unit $i$ is from unit $j$.  In this case \begin{equation}\label{eq:h_within}
        h(\vec{D}, i) = \max_{i \neq j} D_j * 1[ d(i,j) < \bar{d} ] 
    \end{equation}
    This exposure mapping is a realistic specification when it is assumed that spillover effects do not decay over distance until $\bar{d}$ and the intensity of spillovers does not depend on the number of neighboring units treated. For example, this exposure mapping likely applies in the context of new library creation \citep{Berkes_Nencka_2020}. In this case the distance $\bar{d}$ would be the maximum distance people would travel to a nearby library. The benefits of access to a neighboring town library does not depend on whether people can access 1 or more nearby libraries, so the binary exposure mapping is a good approximation to spillovers in this context.  
\end{example}

\begin{example}
    Second, $h(\vec{D}, i)$ can be a function that equals the number of units treated within distance $\bar{d}$, i.e. \begin{equation}\label{eq:h_within_additive}
        h(\vec{D}, i) = \sum_{j} D_j * 1[ d(i,j) < \bar{d} ].
    \end{equation}
    This exposure mapping is no longer binary, so the intensity of spillovers depend on the number of nearby units treated. In the context of large store openings, this exposure mapping captures the additive nature of agglomeration economies. As more nearby counties receive new stores, the agglomeration forces increase (e.g. \citet{Basker_2005}).
\end{example}

\begin{example}
    Last, $h(\vec{D}, i)$ is a spatial decay function where exposure decreases with distance. In this case, spillover intensity is the sum across all treated observations' decay term, i.e. \begin{equation}\label{eq:h_decay}
        h(\vec{D}, i) = \sum_{j \neq i} D_j e^{-\alpha d(i,j)}.
    \end{equation} 
    This exposure mapping allows for the intensity of spillovers to depend on distance to treatment and also is additive in the number of nearby units treated.\footnote{However, this specification assumes that all units are affected by all other units. This creates problems with inference because it implies potential correlation between all units' error terms. For this reason, this function often is summed over only the $k$-nearest neighbors or over units within $\bar{d}$ miles.} The speed that spillover effects decay over distance depends on the parameter $\alpha$ that can be calibrated by the researcher or estimated using a non-linear least squares routine. In the literature on R\&D investment, for example, \citet{Keller_2002} uses a modified version of this exposure mapping where $D_j$ is country $j$'s R\&D expenditure. This specification features exponential decay over distance which captures the theoretical insight that research spending in closer locations has a larger effect than further away spending.
\end{example}

After chosing an exposure mapping that makes sense in the economic context, a researcher must also specify the functional form of the potential outcomes. Typically, the exposure mapping enters into the regression linearly and potentially the coefficient is allowed to differ by treatment status to reflect that spillovers on control and treated units can be different phenomenon.

@@

With the potential outcomes defined, I now formalize what is meant by `spatial spillovers'. I define `spillover onto control units' as: \[
    Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}).
\] 
The spillover measures the difference in non-treated potential outcomes between being exposed at intensity $h(\vec{D}, i)$ and not being exposed. Then, the average spillover effect onto control units averages over potential heterogeneity in the effect size of spillovers and over heterogeneity in exposure intensity $h(\vec{D}, i)$: \[
    \tau_{\text{spill,control}} \equiv \mathbb{E} \left[ Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 0 \right].
\]
To emphasize, the average spillover effect onto conrol units averages over each control unit's exposure mapping. For example, assume the potential outcomes is additively linear in $D_i$ and $h(\vec{D}, i)$ and that the coefficient $\beta_{\text{spill,control}}$ measures the effect of $h(\vec{D}, i)$ on outcome $Y$ among control units. Then an individual control unit's spillover effect is $\beta_{\text{spill,control}} \ h(\vec{D}, i)$. The average spillover effect onto control unit would therefore be $\tau_{\text{spill,control}} = \beta_{\text{spill,control}} * \mathbb{E} \left[ h(\vec{D}, i)\right]$, i.e. the average over all control units exposure mapping. Similarly, we define the average spillover effect onto also treated units as: \[ 
    \tau_{\text{spill,treated}} \equiv \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0) \ \vert \ D_i = 1 \right].
\]

It is important to clarify what I am assuming is the estimand of interest researchers would like to estimate when using difference-in-differences. I assume that what the `average treatment effect' is trying to measure in this context is what I will call the `direct effect of treatment': \[
    \tau_{\text{direct}} = \mathbb{E} \left[ Y_{it}(1, \vec{0}) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 1 \right],
\] 
which measures the effect of being treated in the absence of exposure to spillovers.

My definition of the direct effect of treatment differs from \citet{Sävje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(0, h(\vec{D}, i)) \ \vert \ D_i = 1 \right],
\] 
where the expectation is over individuals and their exposures. The difference between their treatment effect and mine becomes more clear by adding and subtracting terms and rearranging:
\begin{align*}
    &Y_{it}(1, h(\vec{D}, i)) - Y_{it}(0, h(\vec{D}, i)) \\
    &\quad = Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0) + Y_{it}(1, 0) - Y_{it}(0,0) + Y_{it}(0,0) - Y_{it}(0, h(\vec{D}, i)) \\
    &\quad = \underbrace{Y_{it}(1, 0) - Y_{it}(0,0)}_{\text{Direct Effect}} + \underbrace{Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0)}_{\text{Spillover on Treated}} - \underbrace{(Y_{it}(0, h(\vec{D}, i) - Y_{it}(0,0))}_{\text{Spillover on Control}} 
\end{align*}
The individual treatment effect \citet{Sävje_Aronow_Hudgens_2019} use is the direct effect of treatment plus the difference between spillover on treated units and spillover on control units.

I prefer my definition of the direct effect because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. For example, a county deciding whether to implement a policy will want to consider the treatment effect plus the difference in spillover effects at a particular level of exposure. The estimand proposed by \citet{Sävje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at specific exposure intensities. Estimates of the direct effect and spillover effect can be combined to estimate the ATE as defined by \citet{Sävje_Aronow_Hudgens_2019}.

@@

In this section, I identify the two sources of bias in difference-in-differences estimation of the direct effect of treatment. Researchers typically estimate the canonical two-way fixed effects model, 
\begin{equation}\label{eq:twfe}    
    y_{it} = \tau D_{it} + \mu_i + \mu_t + \epsilon_{it},
\end{equation}
where $D_{it} = D_i 1(t=1)$. The estimator $\hat{\tau}$ is a biased estimate for $\tau_{\text{direct}}$ in the presence of spillovers. To show this, I first present the equivalent to the parallel counterfactual trends assumption in the context of the new potential outcome framework.

\begin{assumption}[Parallel Counterfactual Trends]\label{assumption:parallel}
    \[ 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 1 \right] = 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 0 \right]
    \]
\end{assumption}
This assumption states that in the absence of treatment and with zero exposure (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status. This generalizes to the classic parallel counterfactual trends when SUTVA is satisfied because then every unit has zero exposure.

Given that Assumption \ref{assumption:parallel} holds, the estimate $\hat{\tau}$ from (\ref{eq:twfe}) can be decomposed as the direct effect and the two sources of spillover bias. The proof is given in Appendix \ref{sec:proofs}.

\begin{proposition}[Decomposition of Difference-in-Differences Estimate]\label{thm:bias}\ \\    
    If Assumption \ref{assumption:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \underbrace{\mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 1 \right] - \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 0 \right]}_{\text{Difference-in-Differences}} \\ 
        &= 
        \mathbb{E} \left[ Y_{i1}(1, \vec{0}) - Y_{i1}(0, \vec{0}) \mid D_i = 1 \right] + \mathbb{E} \left[ Y_{i1}(1, h(\vec{D}, i)) - Y_{i1}(1, \vec{0}) \mid D_i = 1 \right] \\
        &\quad - \mathbb{E} \left[ Y_{i1}(0, h(\vec{D}, i)) - Y_{i1}(0, \vec{0}) \mid D_i = 0 \right] \\
        &= \tau_{\text{direct}} + \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}
    \end{align*}
\end{proposition}
